%%%% Copyright (C) 2005 -- 2014, Christopher Mark Gore,
%%%% Soli Deo Gloria,
%%%% All rights reserved.
%%%%
%%%% 2317 South River Road, Saint Charles, Missouri 63303 USA.
%%%% Web: http://cgore.com
%%%% Email: cgore@cgore.com
%%%%
%%%% Redistribution and use in source and binary forms, with or without
%%%% modification, are permitted provided that the following conditions are met:
%%%%
%%%%     * Redistributions of source code must retain the above copyright
%%%%       notice, this list of conditions and the following disclaimer.
%%%%
%%%%     * Redistributions in binary form must reproduce the above copyright
%%%%       notice, this list of conditions and the following disclaimer in the
%%%%       documentation and/or other materials provided with the distribution.
%%%%
%%%%     * Neither the name of Christopher Mark Gore nor the names of other
%%%%       contributors may be used to endorse or promote products derived from
%%%%       this software without specific prior written permission.
%%%%
%%%% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
%%%% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
%%%% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
%%%% ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT HOLDER OR CONTRIBUTORS BE
%%%% LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
%%%% CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
%%%% SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
%%%% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
%%%% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
%%%% ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
%%%% POSSIBILITY OF SUCH DAMAGE.

\chapter{Introduction}

The $\Sigma$ library is a generic library of mostly random useful code for
\textsc{Ansi}~Common~Lisp.  It is currently only really focused on
\textsc{Sbcl}, but patches to add support for other systems are more than
welcome.

This library started out as a single file, \texttt{utilities.lisp}, that I
personally used for shared generic code for all of my Lisp code.  Most lispers
have a similar file of some name, \texttt{utilities.lisp}, \texttt{misc.lisp},
\texttt{shared.lisp}, or even \texttt{stuff.lisp}, that is just a random
collection of useful little generic macros and functions.  Mine has grown over
the years, and in 2012 I decided that I should try to make it useful to people
other than myself.

You can download the library from GitHub at:\\
\url{https://github.com/cgore/sigma}\\
and I have some other information on it at my own website at:\\
\url{http://cgore.com/programming/lisp/sigma/}

\section*{Getting Lisp}

Before using this library you need a working Lisp.  I use and recommend
\textsc{Sbcl}, Steel Bank Common Lisp, which is available at:\\
\url{http://www.sbcl.org}\\
This is derived from \textsc{CMUCL}, Carnegie Mellon University Common Lisp,
which is still under active development and is: available at:\\
\url{http://www.cons.org/cmucl/}

\textsc{SBCL} has information on getting started at:\\
\url{http://www.sbcl.org/getting.html}\\
If you are using Debian or a similar Linux distribution (including Ubuntu), you
can just run as root:\\
\texttt{apt-get install sbcl sbcl-doc sbcl-source}

\section*{Getting \textsc{Emacs} and \textsc{Slime}}

After installing, the best way to interact with any Common Lisp is via
\textsc{Slime}, the Superior Lisp Interaction Mode for \textsc{Emacs}, which is
available at:\\
\url{http://common-lisp.net/project/slime/}\\
This can be installed on Debian by:\\
\texttt{apt-get install slime emacs emacs-goodies-el}

\section*{Using the Library}

First we need to clone the utilities.\\
\texttt{mkdir -p ~/programming/lisp}\\
\texttt{cd ~/programming/lisp}\\
\texttt{git clone git@github.com:cgore/sigma.git}

Now we need to make a directory for our project and symlink to the ASDF
definition.  There are other ways to load ASDF libraries, especially if you want
to have them available globally; I strongly recommend you read the documentation
to ASDF.\\
\texttt{mkdir our-new-project}\\
\texttt{cd our-new-project}\\
\texttt{ln -s ~/programming/lisp/sigma/sigma.asd}

Now we need to start up our Lisp REPL.  The best way to do this for perfonal use
is SLIME from within Emacs, but I will demonstrate using the shell itself here.\\
\texttt{sbcl}

Now we are in SBCL.  Let's calculate something.

$$\sum_{i=0}^{100} i$$

\begin{lstlisting}
(require :asdf)} ; Require ASDF
(require :sigma) ; Require the Sigma system via ASDF.
(sigma:use-all-sigma) ; This will pollute COMMON-LISP-USER.
(sum (loop for i from 1 to 100
           collect i)) ; Returns 5050, and makes Euler sad.
\end{lstlisting}

Have fun!
